Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-08-22 paper.qmd
Updated Figure 8

```{r}
#| label: hexbin-regular-scurve1
## hexagon binning to have regular hexagons
hb_obj_scurve1 <- hex_binning(
  data = umap_scurve_scaled, 
  bin1 = 8, 
  r2 = r2)

a1_1 <- calc_bins_y(
  bin1 = 8, 
  r2 = r2
)$a1

@@

## To obtain the standardise counts within hexbins
counts_df1 <- hb_obj_scurve1$std_cts
df_bin_centroids1 <- extract_hexbin_centroids(
  centroids_df = all_centroids_df1, 
  counts_df = counts_df1) |>
  filter(drop_empty == FALSE) |>
  mutate(b1 = "b1 = 8")

@@

```{r}
#| label: hexbin-regular-scurve2
## hexagon binning to have regular hexagons
hb_obj_scurve2 <- hex_binning(
  data = umap_scurve_scaled, 
  bin1 = 10, 
  r2 = r2)

a1_2 <- calc_bins_y(
  bin1 = 10, 
  r2 = r2
  )$a1

@@

## To obtain the standardise counts within hexbins
counts_df2 <- hb_obj_scurve2$std_cts
df_bin_centroids2 <- extract_hexbin_centroids(
  centroids_df = all_centroids_df2, 
  counts_df = counts_df2) |>
  filter(drop_empty == FALSE) |>
  mutate(b1 = "b1 = 10")

@@

```{r}
#| label: hexbin-regular-scurve3
## hexagon binning to have regular hexagons
hb_obj_scurve3 <- hex_binning(
  data = umap_scurve_scaled, 
  bin1 = 14, 
  r2 = r2)

a1_3 <- calc_bins_y(
  bin1 = 14, 
  r2 = r2
)$a1

@@

```{r}
#| echo: false
#| label: fig-bins-scurve
#| fig-pos: H
#| fig-cap: "Hexbin density plots of UMAP layout of the S-curve data, using three different bin inputs: (a) $b = 91 \\text{ } (7, \\text{ }13)  \\text{ }(q = 0.12)$, (b) $b = 220 \\text{ } (11, \\text{ }20) \\text{ }(q = 0.07)$, and (c) $b = 527 \\text{ } (17, \\text{ }31) \\text{ }(q = 0.05)$. Color indicates standardized counts, dark indicating high count and light indicates low count. At the smallest bin size the data segregates into two separate groups, suggesting this is too many bins. Using the MSE of the model fit in $p\\text{-}D$ helps decide on a useful choice of number of bins."
#| fig-width: 6
#| fig-height: 4

>>>>

```{r}
#| label: hexbin-regular-scurve1
## hexagon binning to have regular hexagons
hb_obj_scurve1 <- hex_binning(
  data = umap_scurve_scaled, 
  bin1 = 7, 
  r2 = r2)

a1_1 <- calc_bins_y(
  bin1 = 7, 
  r2 = r2
)$a1

@@

## To obtain the standardise counts within hexbins
counts_df1 <- hb_obj_scurve1$std_cts
df_bin_centroids1 <- extract_hexbin_centroids(
  centroids_df = all_centroids_df1, 
  counts_df = counts_df1) |>
  filter(drop_empty == FALSE) |>
  mutate(b1 = "b1 = 7")

@@

```{r}
#| label: hexbin-regular-scurve2
## hexagon binning to have regular hexagons
hb_obj_scurve2 <- hex_binning(
  data = umap_scurve_scaled, 
  bin1 = 11, 
  r2 = r2)

a1_2 <- calc_bins_y(
  bin1 = 11, 
  r2 = r2
  )$a1

@@

## To obtain the standardise counts within hexbins
counts_df2 <- hb_obj_scurve2$std_cts
df_bin_centroids2 <- extract_hexbin_centroids(
  centroids_df = all_centroids_df2, 
  counts_df = counts_df2) |>
  filter(drop_empty == FALSE) |>
  mutate(b1 = "b1 = 11")

@@

```{r}
#| label: hexbin-regular-scurve3
## hexagon binning to have regular hexagons
hb_obj_scurve3 <- hex_binning(
  data = umap_scurve_scaled, 
  bin1 = 13, 
  r2 = r2)

a1_3 <- calc_bins_y(
  bin1 = 13, 
  r2 = r2
)$a1

@@

```{r}
#| echo: false
#| label: fig-bins-scurve
#| fig-pos: H
#| fig-cap: "Hexbin density plots of UMAP layout of the S-curve data, using three different bin inputs: (a) $b = 91 \\text{ } (7, \\text{ }13)$, (b) $b = 220 \\text{ } (11, \\text{ }20)$, and (c) $b = 312 \\text{ } (13, \\text{ }24)$. Color indicates standardized counts, dark indicating high count and light indicates low count. At the smallest bin size the data segregates into two separate groups, suggesting this is too many bins. Using the MSE of the model fit in $p\\text{-}D$ helps decide on a useful choice of number of bins."
#| fig-width: 6
#| fig-height: 4
